Revisiting `Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation

### 2023-06-28 analysis/article/paper.qmd
pre-submission commit

keywords: 
  - Public goods
  - Dynamic games
  - Reproducibility
  - Generalizability
  - Online Experiment
  - Climate change mitigation
date: last-modified
bibliography:  ../biblio.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Behavioral and Experimental Economics
      formatting: review
      model: 3p
      layout: onecolumn
      cite-style: authoryear
      highlights:
        - Highlight 1
        - Highlight 2 
        - Highlight 3
---

@@

@@

@AGM2018 conducted static public goods games in the lab and on MTurk to draw (and report) lessons from online experimentation. This study extends this literature [see, e.g., @SnowbergYariv_2021; @GuptaRigottiWilson_2021; @GoodmanPaolacci2017; @AmirEtAl2012] by focusing on an _inconvenient_ sample (i.e. one that is a completely inexperienced sample that has not been exposed to interactive experiments before) playing a computationally more complex game. This required me to design robust (and thus, more complex) software to minimize attrition. I collected para data---which capture screen time and survey navigation, for instance---to assess the desired fluency and feasibility of the experiment. Like @AGM2018, this research is of practical relevance as it reports on the robust design that made the dynamic game feasible.

@@

Well before credibility crises comprised a variety of disciplines^[See e.g., @CamererEtAl2016; @BrodeurEtAl2016; @BrodeurEtAl2020; @ChristensenMiguel2018; @FerraroShukla2020 who focus on economics.], @bemwriting [p. 2] indirectly conceded discrepancies between fuzzy research processes and the polished article that results from it:

@@

Now, 20 years later, we accustom ourselves to mechanisms designed to unravel the fuzzy back-and-forth between exploratory and confirmatory research. Prominently, pre-registrations were established to tie the researchers' hands with respect to p-hacking and ex-post theorizing when doing confirmatory research---reducing observable discrepancies between Daryl Bem's above-mentioned (a) and (b).

While pre-registrations could help to make empirical sciences more credible, they come at costs that are discussed by @Olken2015 as well as @Krishna2021, @SimmonsEtAl2021, and @PhamEtAl2021. One important concern is that strict pre-registrations may severely discount any additional analysis of the data inspired by surprise results. In practice, there are at least two ways to approach that problem: First, one could (and should) add unforeseen analyses to the article but label them as exploratory. Second, one could write a minimal pre-registration leaving enough degrees of freedom for the researcher to incorporate such surprises without further notice. The key difference between the two approaches is the deliberate omission of transparency.

@@

This erosion of credibility can be avoided by the means of transparency---not because transparency can tighten the ties but also because it facilitates scrutiny, policing [see @AnkelPetersFialaNeubauer2023] and forensics: transparent research practices make it easier for others to track and understand a researcher's reaction to surprises in the data, which, in turn, gives the researcher more flexibility to react to these surprises. As additional transparency is associated with additional effort, I propose version control as a natural and easy-to-implement solution augmenting pre-registrations.

@@

The experiment reported in this article qualifies for the _'pre-registered'_ label at first glance:
The analysis was pre-registered in the American Economic Association's RCT Registry [@preregistration]. Further, I pre-registered the exact analyses I planned to run when I designed the experiment on [GitHub](https://github.com/Howquez/GMTV/tree/July21Replication/analysis/R).^[bit.ly/3Pnak5H. To run the code, you need to execute the .Rmd files in this repository in the order that is indicated by file names. Detailed instructions can be found in the README file accompanying the analysis scripts.]

@@

The experiment was logistically complex for several reasons. First, the sample was inexperienced. Second, the experiment was interactive and synchronous. Third, the underlying game was dynamic and interdependent. This makes dropouts not only more likely but also more expensive, which is why attrition was a major concern in implementing the experiment.

@@

Turning to the time spent on each page, I focus on the decision times in the dynamic public goods game as @Anderhub2001 did. How many seconds did the participants need to decide in each period of the game? Not too many. @fig-time-spent illustrates an intuitive pattern: The first decision took about `r tmp[round == 1, round(time_spent)]` seconds. The second decision---where participants first learned about the other group members' previous decisions---took longer (about `r tmp[round == 2, round(time_spent)]` seconds). Subsequently, decision times first declined and stabilized at `r tmp[round > 3, round(mean(time_spent))]` seconds. Importantly, decision times were so short that crosstalk, that is, communication through private channels---a common concern^[See, for instance, the discussion section in @AGM2018 [p. 119].] in online experiments---was unlikely, especially because it would require the identification of other group members.^[There were only `r plotDT[page_index == 12 & time_spent > 60, .N]` participants (from all four sessions) who needed more than 60 seconds to make the second decision.]

@@

I gratefully acknowledge support by the German Research Foundation (DFG) under Germany’s Excellence Strategy, cluster EXC 2037 “Climate, Climatic Change, and Society” (project 390683824). I thank Stefan Traub and the participants of the CLICCS B5 group for helpful discussions.

>>>>

keywords: 
  - Public goods
  - Dynamic games
  - Reproducibility
  - Generalizability
  - Online Experiment
  - Climate change mitigation
date: last-modified
bibliography:  ../biblio.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Behavioral and Experimental Economics
      formatting: review
      model: 3p
      layout: onecolumn
      cite-style: authoryear
---

@@

# Importantly, as of today, I don't remember how to perfectly identify students in the last session. Hence, I label all participants in that session as students. This is reasonable when analyzing group level outcomes because it is likely that there was at least one student in many of the groups. It is somewhat problematic in the generalization discussion. However, using age as an additional variable to guess student status (e.g. non-student if age > 30) does not not change the results qualitatively.

@@

@AGM2018 conducted static public goods games in the lab and on MTurk to draw (and report) lessons from online experimentation. This study extends this literature [see, e.g., @KrantzEtAl1997, @AmirEtAl2012; @KleinEtAl2014; @PaolacciChandler2014; @GoodmanPaolacci2017; @SnowbergYariv_2021; @GuptaRigottiWilson_2021; @BusoEtAl2021] by focusing on an _inconvenient_ sample (i.e. one that is a completely inexperienced sample [see also @BenndorfEtAl2017] that has not been exposed to interactive experiments before) playing a computationally more complex game during a time that was characterized by more inattentiveness in online samples [@ArecharRand2021; @PeytonEtAl2022]. This required me to design robust (and thus, more complex) software to minimize attrition. I collected para data [@Parsons2022 p. 12]---which capture screen time and survey navigation, for instance---to assess the desired fluency and feasibility of the experiment. Like @AGM2018, this research is of practical relevance as it reports on the robust design that made the dynamic game feasible.

@@

Well before credibility crises comprised a variety of disciplines^[See e.g., @CamererEtAl2016; @BrodeurEtAl2016; @BrodeurEtAl2020; @ChristensenMiguel2018; @FerraroShukla2020; @PageEtAl2021 who focus on economics.], @bemwriting [p. 2] indirectly conceded discrepancies between fuzzy research processes and the polished article that results from it:

@@

Now, 20 years later, we accustom ourselves to mechanisms designed to unravel the fuzzy back-and-forth between exploratory and confirmatory research. Prominently, pre-registrations and pre-analysis plans (PAP) [see, e.g., @BrodeurEtAl2022] were established to tie the researchers' hands with respect to p-hacking and ex-post theorizing when doing confirmatory research---reducing observable discrepancies between Daryl Bem's above-mentioned (a) and (b).

While pre-registrations and PAPs could help to make empirical sciences more credible, they come at costs that are discussed by @Olken2015, @PageEtAl2021, as well as @Krishna2021, @SimmonsEtAl2021, and @PhamEtAl2021. One important concern is that strict pre-registrations may severely discount any additional analysis of the data inspired by surprise results. In practice, there are at least two ways to approach that problem: First, one could (and should) add unforeseen analyses to the article but label them as exploratory. Second, one could write a minimal pre-registration leaving enough degrees of freedom for the researcher to incorporate such surprises without further notice. The key difference between the two approaches is the deliberate omission of transparency.

@@

This erosion of credibility can be avoided by the means of transparency---not because transparency can tighten the ties but because it facilitates scrutiny, policing [see @AnkelPetersFialaNeubauer2023] and forensics [see, e.g., @SimonsohnEtAl2023]: transparent research practices make it easier for others to track and understand a researcher's reaction to surprises in the data, which, in turn, gives the researcher more flexibility to react to these surprises. As additional transparency is associated with additional effort, I propose version control as a natural and easy-to-implement solution augmenting pre-registrations.

@@

The experiment reported in this article qualifies for the _'pre-registered'_ label at first glance:
The analysis was pre-registered in the American Economic Association's RCT Registry [@preregistration]. Further, I pre-registered the exact analyses I planned to run (PAP) when I designed the experiment on [GitHub](https://github.com/Howquez/GMTV/tree/July21Replication/analysis/R).^[bit.ly/3Pnak5H. To run the code, you need to execute the .Rmd files in this repository in the order that is indicated by file names. Detailed instructions can be found in the README file accompanying the analysis scripts.]

@@

The experiment was logistically complex for several reasons. First, the sample was inexperienced. Second, the experiment was interactive and synchronous. Third, the underlying game was dynamic and interdependent. This makes dropouts not only more likely but also more expensive, which is why attrition  was a major concern in implementing the experiment.^[see @ZhouFischbach2016 for a nice illustration of how (selective) attrition affects identification.]

@@

Turning to the time spent on each page, I focus on the decision times in the dynamic public goods game as @Anderhub2001 did. How many seconds did the participants need to decide in each period of the game? Not too many. @fig-time-spent illustrates an plausible pattern: The first decision took about `r tmp[round == 1, round(time_spent)]` seconds. The second decision---where participants first learned about the other group members' previous decisions---took longer (about `r tmp[round == 2, round(time_spent)]` seconds). Subsequently, decision times first declined and stabilized at `r tmp[round > 3, round(mean(time_spent))]` seconds. Importantly, decision times were so short that crosstalk, that is, communication through private channels---a common concern^[See, for instance, the discussion section in @AGM2018 [p. 119].] in online experiments---was unlikely, especially because it would require the identification of other group members.^[There were only `r plotDT[page_index == 12 & time_spent > 60, .N]` participants (from all four sessions) who needed more than 60 seconds to make the second decision.]

@@

I gratefully acknowledge support by the German Research Foundation (DFG) under Germany’s Excellence Strategy, cluster EXC 2037 “Climate, Climatic Change, and Society” (project 390683824). I thank Stefan Traub, whoever proofreads this and the participants of the CLICCS B5 group for helpful discussions.

### 2023-06-28 analysis/article/paper.tex
pre-submission commit

\begin{highlights}
\item Highlight 1\item Highlight 2\item Highlight 3
\end{highlights}

\begin{keyword}
    Public goods \sep Dynamic
games \sep Reproducibility \sep Generalizability \sep Online
Experiment \sep 
    Climate change mitigation
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

\citet{AGM2018} conducted static public goods games in the lab and on
MTurk to draw (and report) lessons from online experimentation. This
study extends this literature \citep[see,
e.g.,][]{SnowbergYariv_2021, GuptaRigottiWilson_2021, GoodmanPaolacci2017, AmirEtAl2012}
by focusing on an \emph{inconvenient} sample (i.e.~one that is a
completely inexperienced sample that has not been exposed to interactive
experiments before) playing a computationally more complex game. This
required me to design robust (and thus, more complex) software to
minimize attrition. I collected para data---which capture screen time
and survey navigation, for instance---to assess the desired fluency and
feasibility of the experiment. Like \citet{AGM2018}, this research is of
practical relevance as it reports on the robust design that made the
dynamic game feasible.

@@

Well before credibility crises comprised a variety of
disciplines\footnote{See e.g., \citet{CamererEtAl2016};
  \citet{BrodeurEtAl2016}; \citet{BrodeurEtAl2020};
  \citet{ChristensenMiguel2018}; \citet{FerraroShukla2020} who focus on
  economics.}, \citet[p.~2]{bemwriting} indirectly conceded
discrepancies between fuzzy research processes and the polished article
that results from it:

@@

Now, 20 years later, we accustom ourselves to mechanisms designed to
unravel the fuzzy back-and-forth between exploratory and confirmatory
research. Prominently, pre-registrations were established to tie the
researchers' hands with respect to p-hacking and ex-post theorizing when
doing confirmatory research---reducing observable discrepancies between
Daryl Bem's above-mentioned (a) and (b).

While pre-registrations could help to make empirical sciences more
credible, they come at costs that are discussed by \citet{Olken2015} as
well as \citet{Krishna2021}, \citet{SimmonsEtAl2021}, and
\citet{PhamEtAl2021}. One important concern is that strict
pre-registrations may severely discount any additional analysis of the
data inspired by surprise results. In practice, there are at least two
ways to approach that problem: First, one could (and should) add
unforeseen analyses to the article but label them as exploratory.
Second, one could write a minimal pre-registration leaving enough
degrees of freedom for the researcher to incorporate such surprises
without further notice. The key difference between the two approaches is
the deliberate omission of transparency.

@@

This erosion of credibility can be avoided by the means of
transparency---not because transparency can tighten the ties but also
because it facilitates scrutiny, policing
\citep[see][]{AnkelPetersFialaNeubauer2023} and forensics: transparent
research practices make it easier for others to track and understand a
researcher's reaction to surprises in the data, which, in turn, gives
the researcher more flexibility to react to these surprises. As
additional transparency is associated with additional effort, I propose
version control as a natural and easy-to-implement solution augmenting
pre-registrations.

@@

The experiment reported in this article qualifies for the
\emph{`pre-registered'} label at first glance: The analysis was
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. Further, I pre-registered the exact analyses I
planned to run when I designed the experiment on
\href{https://github.com/Howquez/GMTV/tree/July21Replication/analysis/R}{GitHub}.\footnote{bit.ly/3Pnak5H.
  To run the code, you need to execute the .Rmd files in this repository
  in the order that is indicated by file names. Detailed instructions
  can be found in the README file accompanying the analysis scripts.}

@@

The experiment was logistically complex for several reasons. First, the
sample was inexperienced. Second, the experiment was interactive and
synchronous. Third, the underlying game was dynamic and interdependent.
This makes dropouts not only more likely but also more expensive, which
is why attrition was a major concern in implementing the experiment.

@@

Turning to the time spent on each page, I focus on the decision times in
the dynamic public goods game as \citet{Anderhub2001} did. How many
seconds did the participants need to decide in each period of the game?
Not too many. Figure~\ref{fig-time-spent} illustrates an intuitive
pattern: The first decision took about 22 seconds. The second
decision---where participants first learned about the other group
members' previous decisions---took longer (about 33 seconds).
Subsequently, decision times first declined and stabilized at 19
seconds. Importantly, decision times were so short that crosstalk, that
is, communication through private channels---a common concern\footnote{See,
  for instance, the discussion section in \citet[p.~119]{AGM2018}.} in
online experiments---was unlikely, especially because it would require
the identification of other group members.\footnote{There were only 9
  participants (from all four sessions) who needed more than 60 seconds
  to make the second decision.}

@@

I gratefully acknowledge support by the German Research Foundation (DFG)
under Germany's Excellence Strategy, cluster EXC 2037 ``Climate,
Climatic Change, and Society'' (project 390683824). I thank Stefan Traub
and the participants of the CLICCS B5 group for helpful discussions.

>>>>

\begin{keyword}
    Public goods \sep Dynamic
games \sep Reproducibility \sep Generalizability \sep Online
Experiment \sep 
    Climate change mitigation
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\citet{AGM2018} conducted static public goods games in the lab and on
MTurk to draw (and report) lessons from online experimentation. This
study extends this literature \citetext{\citealp[see,
e.g.,][\citet{AmirEtAl2012}]{KrantzEtAl1997}; \citealp{KleinEtAl2014}; \citealp{PaolacciChandler2014}; \citealp{GoodmanPaolacci2017}; \citealp{SnowbergYariv_2021}; \citealp{GuptaRigottiWilson_2021}; \citealp{BusoEtAl2021}}
by focusing on an \emph{inconvenient} sample (i.e.~one that is a
completely inexperienced sample \citep[see also][]{BenndorfEtAl2017}
that has not been exposed to interactive experiments before) playing a
computationally more complex game during a time that was characterized
by more inattentiveness in online samples
\citep{ArecharRand2021, PeytonEtAl2022}. This required me to design
robust (and thus, more complex) software to minimize attrition. I
collected para data \citep[ p.~12]{Parsons2022}---which capture screen
time and survey navigation, for instance---to assess the desired fluency
and feasibility of the experiment. Like \citet{AGM2018}, this research
is of practical relevance as it reports on the robust design that made
the dynamic game feasible.

@@

Well before credibility crises comprised a variety of
disciplines\footnote{See e.g., \citet{CamererEtAl2016};
  \citet{BrodeurEtAl2016}; \citet{BrodeurEtAl2020};
  \citet{ChristensenMiguel2018}; \citet{FerraroShukla2020};
  \citet{PageEtAl2021} who focus on economics.},
\citet[p.~2]{bemwriting} indirectly conceded discrepancies between fuzzy
research processes and the polished article that results from it:

@@

Now, 20 years later, we accustom ourselves to mechanisms designed to
unravel the fuzzy back-and-forth between exploratory and confirmatory
research. Prominently, pre-registrations and pre-analysis plans (PAP)
\citep[see, e.g.,][]{BrodeurEtAl2022} were established to tie the
researchers' hands with respect to p-hacking and ex-post theorizing when
doing confirmatory research---reducing observable discrepancies between
Daryl Bem's above-mentioned (a) and (b).

While pre-registrations and PAPs could help to make empirical sciences
more credible, they come at costs that are discussed by
\citet{Olken2015}, \citet{PageEtAl2021}, as well as \citet{Krishna2021},
\citet{SimmonsEtAl2021}, and \citet{PhamEtAl2021}. One important concern
is that strict pre-registrations may severely discount any additional
analysis of the data inspired by surprise results. In practice, there
are at least two ways to approach that problem: First, one could (and
should) add unforeseen analyses to the article but label them as
exploratory. Second, one could write a minimal pre-registration leaving
enough degrees of freedom for the researcher to incorporate such
surprises without further notice. The key difference between the two
approaches is the deliberate omission of transparency.

@@

This erosion of credibility can be avoided by the means of
transparency---not because transparency can tighten the ties but because
it facilitates scrutiny, policing
\citep[see][]{AnkelPetersFialaNeubauer2023} and forensics \citep[see,
e.g.,][]{SimonsohnEtAl2023}: transparent research practices make it
easier for others to track and understand a researcher's reaction to
surprises in the data, which, in turn, gives the researcher more
flexibility to react to these surprises. As additional transparency is
associated with additional effort, I propose version control as a
natural and easy-to-implement solution augmenting pre-registrations.

@@

The experiment reported in this article qualifies for the
\emph{`pre-registered'} label at first glance: The analysis was
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. Further, I pre-registered the exact analyses I
planned to run (PAP) when I designed the experiment on
\href{https://github.com/Howquez/GMTV/tree/July21Replication/analysis/R}{GitHub}.\footnote{bit.ly/3Pnak5H.
  To run the code, you need to execute the .Rmd files in this repository
  in the order that is indicated by file names. Detailed instructions
  can be found in the README file accompanying the analysis scripts.}

@@

The experiment was logistically complex for several reasons. First, the
sample was inexperienced. Second, the experiment was interactive and
synchronous. Third, the underlying game was dynamic and interdependent.
This makes dropouts not only more likely but also more expensive, which
is why attrition was a major concern in implementing the
experiment.\footnote{see \citet{ZhouFischbach2016} for a nice
  illustration of how (selective) attrition affects identification.}

@@

Turning to the time spent on each page, I focus on the decision times in
the dynamic public goods game as \citet{Anderhub2001} did. How many
seconds did the participants need to decide in each period of the game?
Not too many. Figure~\ref{fig-time-spent} illustrates an plausible
pattern: The first decision took about 22 seconds. The second
decision---where participants first learned about the other group
members' previous decisions---took longer (about 33 seconds).
Subsequently, decision times first declined and stabilized at 19
seconds. Importantly, decision times were so short that crosstalk, that
is, communication through private channels---a common concern\footnote{See,
  for instance, the discussion section in \citet[p.~119]{AGM2018}.} in
online experiments---was unlikely, especially because it would require
the identification of other group members.\footnote{There were only 9
  participants (from all four sessions) who needed more than 60 seconds
  to make the second decision.}

@@

I gratefully acknowledge support by the German Research Foundation (DFG)
under Germany's Excellence Strategy, cluster EXC 2037 ``Climate,
Climatic Change, and Society'' (project 390683824). I thank Stefan
Traub, whoever proofreads this and the participants of the CLICCS B5
group for helpful discussions.
